ESRC SDAI Application

### 2019-07-03 case_for_support.tex
Anna update

The first dataset is provided by the West Midlands Police, and includes all recorded crimes and incidents from the period between January, 2010 and October, 2018. The West Midlands Police is the third largest police force in England, serving a population of 2.9 million \textit{(ref)}. Crimes that have a domestic abuse marker indicate cases of domestic abuse that meet the criteria for notifiable offences in the UK, whereas domestic abuse incidents refer to cases that do not qualify as a crime. About 31\% of all crimes and incidents have a domestic abuse marker. For each record in this dataset, we have information about the exact location and time of the incident or crime, the gender, age and ethnicity of the offender and victim, and the severity of the injury sustained by the victim, if any. The first and last occurrence of the offence, as well as the exact time of reporting is also recorded. Each person in this dataset has a unique person identifier, allowing us to follow people over time. \textit{Sentence about access to this dataset.}

The second dataset is the Crime Survey for England and Wales (CSEW), which is an annual, cross-sectional representative survey collecting information on victims of crime across England and Wales, including 30,000--40,000 households every year \textit{(ref)}. The survey has a self-completion module containing questions about the respondent's experiences of domestic abuse throughout their life (since the age of 16), and in the past 12 months in particular \textit{(ref)}. Given the serious problem of underreporting uniquely characteristic to domestic abuse, the CSEW is the most reliable source of information on the prevalence of domestic abuse in England and Wales.   since when?

@@

 \textcolor{red}{\textit{Short intro about aims, highlighting added value.}}
In this research project, we will first use the CSEW to investigate the demographic and socio-economic characteristics of survivors of domestic abuse. Exploring the risk factors in domestic abuse victimisation is key in designing effectively targeted policy measures. We will also use the WMP data to explore the extent of police mis-recording of domestic abuse. Increasing victim's trust in the police through improving law enforcement response to reported incidents is key in encouraging victims to report domestic abuse and prevent further harm.

@@

\textcolor{red}{\textit{Detailed plan}} First, using the CSEW data, we will extend our understanding of the demographic and socio-economic factors predicting domestic abuse victimisation. The CSEW provides a broad range of information on the characteristics of the respondent, including age, sex, marital status, number of children, ethnicity, education, employment, income, benefit history, physical and mental health, frequency of going out, house and car ownership, self-reported well-being, and frequency of drug and alcohol use. Previous research has identified a number of risk factors predicting domestic abuse victimisation (being female, young, unemployed, separated or divorced, living in single parent household and earning in the lowest income bracket), based on \href{https://www.ons.gov.uk/peoplepopulationandcommunity/crimeandjustice/articles/domesticabusefindingsfromthecrimesurveyforenglandandwales/yearendingmarch2018}{descriptive statistics} derived from the CSEW data. Building on what we know so far, we will explore the predictive power of a broader range of victim characteristics in statistical model, using a logistic regression approach. Investigating the explanatory power of these characteristics within one statistical model will provide us with a deeper understanding of the importance of these factors in predicting domestic abuse victimisation, compared to a purely descriptive statistics approach. We are also interested in how the predictive power of victim characteristics vary with the type of abuse suffered (physical abuse, threats, sexual abuse).

@@

 Police forces in England, and particularly the West Midlands Police, have previously been criticised for mis-recording violent cases of violent crime, including \href{https://www.bbc.co.uk/news/uk-england-468676578}{domestic abuse.} Mis-recording these already heavily underreported crimes decrease victim's trust in the Criminal Justice System and their willingness to report subsequent incidents. \textit{Why don't we use the aggregate number of \href{https://www.ons.gov.uk/peoplepopulationandcommunity/crimeandjustice/datasets/domesticabuseinenglandandwalesappendixtables}{police-recorded domestic abuse cases for each police force} to do this? But if we don't particularly rely on the WMP data is there a point in this? They use \href{https://www.ons.gov.uk/peoplepopulationandcommunity/crimeandjustice/bulletins/domesticabuseinenglandandwales/yearendingmarch2018}{CSEW and police data} in this report, but they don't seem to compare the two}

\begin{table}[!htbp]
\caption{Understanding domestic abuse: survivor characteristics and police mis-recording, regressions}
  \begin{threeparttable}[t]
  \centering
       \begin{tabular}{ L{2.8cm}  C{1.8cm}  C{1.8cm}  C{3.2cm}  C{2.8cm}  C{2cm} }
    \toprule
     \textbf{Research question} & \textbf{Dataset}    & \textbf{Unit of analysis} & \textbf{Outcome variables} & \textbf{Explanatory variables} & \textbf{Model} \\
    \midrule
    What are the characteristics of domestic abuse victims? & CSEW & Individual (respondent-level) & Domestic abuse victimisation; Victimisation by type (physical, sexual abuse, threats) & demographic and socio-economic characteristics & Logistic regression/random forest \\
        \midrule
    What is the extent of police mis-recording of domestic abuse? & ONS data & & & & \\
%    Beznau II & 23.10.1971 & Pressurised water reactor & 365   & 2031 \\
%    Mühleberg & 01.07.1971 & Boiling water reactor & 373   & 2019\\
%    Gösgen & 02.02.1979 & Pressurised water reactor & 1010  & 2039 \\
%    Leibstadt & 24.05.1984 & Boiling water reactor & 1220  & 2044 \\
     \bottomrule
  \end{tabular}
%     \begin{tablenotes}
%     \item[1] Assumption: life time of 60 years.
%     \item[2] Official shut-down.
%   \end{tablenotes}
    \end{threeparttable}%
  \label{tab:addlabel}%
\end{table}%

@@

The first extensive investigation of the demographic and socio-economic predictors of domestic abuse victimisation.

@@

\textcolor{red}{\textit{Detailed plan}} Drawing on the relative strengths of the CSEW and the WMP datasets, we will explore the relative importance of the risk factors predicting serious harm. Previous studies investigating this question mostly relied on police data and focused on \href{https://link.springer.com/article/10.1007/s41887-017-0011-1}{identifying high-risk offenders}. We will complement and extend this research by simultaneously considering the victim characteristics that are predictive of suffering serious injuries using the CSEW. Specifically, using a multinomial logistic approach, we will explore the extent to which financial independence and resilience (income bracket), mental health (self-reported well-being), drug and alcohol dependence (frequency of usage), social isolation (self-reported frequency of going out, living alone or with the perpetrator, family disputes), recent separation, and feeling frightened (increased home security, reporting feeling scared) are predictors of various levels of harm (threats, minor or serious injury) \textit{causality issue?}. We will know the level of harm suffered, because the CSEW asks whether the respondent needed to see a doctor, nurse or other health worker, or needed to take time off work because of the abuse. Given the cross-sectional nature of the CSEW data, this analysis will be exploratory rather than causal.

@@

We will complement this analysis using the WMP, by estimating how living in a household with domestic and child abuse affects the likelihood of engaging in various types of criminal behaviours (property-related vs violent crime). We will do so by identifying what percentage of young offenders are victims of child abuse, and have a home address that is associated with domestic abuse incidents. Given the problem of underreporting, this will give us a lower bound for the estimate of the proportion of children with behavioural problems who live in abusive households. \textit{Am I missing something here?}

\begin{table}[!htbp]
\caption{The long-lasting effects of domestic abuse, regressions}
  \begin{threeparttable}[t]
  \centering
       \begin{tabular}{ L{2.8cm}  C{1.8cm}  C{1.8cm}  C{3.2cm}  C{2.8cm}  C{2cm} }
    \toprule
     \textbf{Research question} & \textbf{Dataset}    & \textbf{Unit of analysis} & \textbf{Outcome variables} & \textbf{Explanatory variables} & \textbf{Model} \\
    \midrule
    How does past childhood abuse affect socio-economic outcomes in adulthood? & CSEW & Individual (respondent-level) & Educational attainment, employment status and health-related outcomes, including domestic abuse victimisation & Degree and type of childhood abuse suffered & Propensity score matching \\
                \midrule
                   How does experiences of past (but not recent) domestic abuse affect socio-economic outcomes in adulthood? & CSEW & Individual (respondent-level) & Current health and employment outcomes & Degree and type of domestic abuse suffered & Propensity score matching \\
                \midrule
     How does experiences of domestic abuse at home and experiences of child abuse affect behavioural outcomes in childhood? & CSEW & Individual (respondent-level) & Bullying, involvement with gangs, school truancy, carrying knives, drugs and alcohol usage, learning difficulties, health & Degree and type of domestic abuse witnessed/child abuse suffered & Propensity score matching \\
                \midrule
          What proportion of young offenders come from abusive households? & WMP & Young offenders &  &  &  \\
     \bottomrule
  \end{tabular}
    \end{threeparttable}%
  \label{tab:addlabel}%
\end{table}%

>>>>

The first dataset is provided by the West Midlands Police, and includes all recorded crimes and incidents from the period between January, 2010 and October, 2018. The West Midlands Police is the third largest police force in England, serving a population of 2.9 million \cite{Homeoffice}. Crimes that have a domestic abuse marker indicate cases of domestic abuse that meet the criteria for notifiable offences in the UK, whereas domestic abuse incidents refer to cases that do not qualify as a crime. About 31\% of all crimes and incidents have a domestic abuse marker. For each record in this dataset, we have information about the exact location and time of the incident or crime, the gender, age and ethnicity of the offender and victim, and the severity of the injury sustained by the victim, if any. The first and last occurrence of the offence, as well as the exact time of reporting is also recorded. Each person in this dataset has a unique person identifier, allowing us to follow people over time. \textit{Sentence about access to this dataset.}

The second dataset is the Crime Survey for England and Wales (CSEW), which is an annual, cross-sectional representative survey collecting information on victims of crime across England and Wales, including 30,000--40,000 households every year \cite{OfficeforNationalStatistics2019}. The survey has a self-completion module containing questions about the respondent's experiences of domestic abuse throughout their life (since the age of 16), and in the past 12 months in particular. Given the serious problem of underreporting uniquely characteristic to domestic abuse, the CSEW is the most reliable source of information on the prevalence of domestic abuse in England and Wales.

@@

 \textcolor{red}{\textit{Short intro about aims, highlighting added value.}}
In this research project, we will first use the CSEW to investigate the demographic and socio-economic characteristics of survivors of domestic abuse. Exploring the risk factors in domestic abuse victimisation is key in designing effectively targeted policy measures. We will also use the CSEW and WMP data to explore the police's response to reported cases of domestic abuse. Increasing victim's trust in the police through improving law enforcement response to reported incidents is key in encouraging victims to report domestic abuse and prevent further harm.

@@

\textcolor{red}{\textit{Detailed plan}} First, using the CSEW data, we will extend our understanding of the demographic and socio-economic factors predicting domestic abuse victimisation. The CSEW provides a broad range of information on the characteristics of the respondent, including age, sex, marital status, number of children, ethnicity, education, employment, income, benefit history, physical and mental health, frequency of going out, house and car ownership, self-reported well-being, and frequency of drug and alcohol use. Previous research has identified a number of risk factors predicting domestic abuse victimisation (being female, young, unemployed, separated or divorced, living in single parent household and earning in the lowest income bracket), based on descriptive statistics derived from the CSEW data \cite{ONS}. Building on what we know so far, we will explore the predictive power of a broader range of victim characteristics in statistical model, using a logistic regression approach. Investigating the explanatory power of these characteristics within one statistical model will provide us with a deeper understanding of the importance of these factors in predicting domestic abuse victimisation, compared to a purely descriptive statistics approach. We are also interested in how the predictive power of victim characteristics vary with the type of abuse suffered (physical abuse, threats, sexual abuse).

@@

 Her Majesty's Inspectorate of Constabulary and Fires \& Rescue Services (HMICFRS) oversees police forces in England Wales, and regularly inspects their responses to reported crime, including cases of domestic abuse. Their latest crime data integrity inspection of 32 police forces have found that 20 forces require significant improvements in their crime recording practices \cite{HerMajestysInspectorateofConstabularyandFires&RescueServices2019}. These concerns are especially pertinent in the context of domestic abuse, for example, the latest report found that the WMP have failed to record 25\% of domestic abuse cases reported to the police, a majority of which would have qualified as a crime. To explore some of the reasons underlying this phenomenon as well as understanding its implications with regards to victims' trust in the CJS, we will limit our sample to those who have reported a specific type of domestic abuse incident (sexual, physical violence, harassment and stalking) in the past 12 months, according to the CSEW. First, we will run a multinomial logistic regression model with police force fixed effects, aiming to explain how the police action taken depend on the socio-economic characteristics of the victim. \textit{For example, it is possible that police forces are less likely to record a case if the victim has a substance dependency. Too speculative?} To complement this analysis and explore potential interactions, we will also investigate this question with a mixed effects random forest with police force clustering. Second, we will also investigate how victim's satisfaction with the police, and their perception of their personal security depends on the police action taken, and their socio-economic characteristics, using a logistic and a random forest statistical framework.

% Using the CSEW, we are interested in understanding how for each offence type (sexual, physical with injury, stalking and harassment), the police's response depend on the victim's socio-economic characteristics 
% 
% From the CSEW, we can estimate what percentage of reported cases that ended in warning, arrest, charge, or no action, a well as the victim's satisfaction with the police handling the case and their perception of their own security as a result of the outcome.  
% 
%https://www.justiceinspectorates.gov.uk/hmicfrs/our-work/article/crime-data-integrity/reports-rolling-programme-crime-data-integrity/

% Police forces in England, and particularly the West Midlands Police, have previously been criticised for mis-recording violent cases of violent crime, including \href{https://www.bbc.co.uk/news/uk-england-468676578}{domestic abuse.} Mis-recording these already heavily underreported crimes decrease victim's trust in the Criminal Justice System and their willingness to report subsequent incidents. \textit{Why don't we use the aggregate number of \href{https://www.ons.gov.uk/peoplepopulationandcommunity/crimeandjustice/datasets/domesticabuseinenglandandwalesappendixtables}{police-recorded domestic abuse cases for each police force} to do this? But if we don't particularly rely on the WMP data is there a point in this? They use \href{https://www.ons.gov.uk/peoplepopulationandcommunity/crimeandjustice/bulletins/domesticabuseinenglandandwales/yearendingmarch2018}{CSEW and police data} in this report, but they don't seem to compare the two}

\begin{table}[!htbp]
\caption{Understanding domestic abuse: survivor characteristics and police mis-recording, regressions}
  \begin{threeparttable}[t]
  \centering
       \begin{tabular}{ L{2.8cm}  C{1.8cm}  C{1.8cm}  C{3.2cm}  C{2.8cm}  C{2cm} }
    \toprule
     \textbf{Research question} & \textbf{Dataset}    & \textbf{Unit of analysis} & \textbf{Outcome variables} & \textbf{Explanatory variables} & \textbf{Model} \\
    \midrule
    What are the characteristics of domestic abuse victims? & CSEW & Individual (respondent-level) & Domestic abuse victimisation; Victimisation by type (physical, sexual abuse, threats) & demographic and socio-economic characteristics & Logistic regression/random forest \\
        \midrule
    How does the police response depend on victim characteristics? & CSEW (reported domestic abuse, by type of offence) & Individual (respondent-level) & Police action taken (no action, warning, arrest, charge) & demographic and socio-economic characteristics of the victim & Multinomial logistic regression with  fixed effects/ mixed effects random forest \\
     \midrule
    How do victims of domestic abuse who reported perceive the police? & CSEW (reported domestic abuse, by type of offence) & Individual (respondent-level)& Satisfaction with police action, perception of security after reporting & police action taken, demographic and socio-economic characteristics of the victim & Multinomial logistic regression with fixed effects/ mixed effects random forest \\
%    Beznau II & 23.10.1971 & Pressurised water reactor & 365   & 2031 \\
%    Mühleberg & 01.07.1971 & Boiling water reactor & 373   & 2019\\
%    Gösgen & 02.02.1979 & Pressurised water reactor & 1010  & 2039 \\
%    Leibstadt & 24.05.1984 & Boiling water reactor & 1220  & 2044 \\
     \bottomrule
  \end{tabular}
%     \begin{tablenotes}
%     \item[1] Assumption: life time of 60 years.
%     \item[2] Official shut-down.
%   \end{tablenotes}
    \end{threeparttable}%
  \label{tab:addlabel}%
\end{table}%

@@

The first extensive investigation of the demographic and socio-economic predictors of domestic abuse victimisation. An exploration of how police response to reported cases, victim satisfaction with the police and perception of security after reporting domestic abuse depend on victim characteristics.

@@

\textcolor{red}{\textit{Detailed plan}} Drawing on the relative strengths of the CSEW and the WMP datasets, we will explore the relative importance of the risk factors predicting serious harm. Previous studies investigating this question mostly relied on police data and focused on identifying high-risk offenders \cite{thornton}. We will complement and extend this research by simultaneously considering the victim characteristics that are predictive of suffering serious injuries using the CSEW. Specifically, using a multinomial logistic approach, we will explore the extent to which financial independence and resilience (income bracket), mental health (self-reported well-being), drug and alcohol dependence (frequency of usage), social isolation (self-reported frequency of going out, living alone or with the perpetrator, family disputes), recent separation, and feeling frightened (increased home security, reporting feeling scared) are predictors of various levels of harm (threats, minor or serious injury) \textit{causality issue?}. We will know the level of harm suffered, because the CSEW asks whether the respondent needed to see a doctor, nurse or other health worker, or needed to take time off work because of the abuse. Given the cross-sectional nature of the CSEW data, this analysis will be exploratory rather than causal.

@@

We will complement this analysis using the WMP, by estimating how living in a household with domestic and child abuse affects the likelihood of engaging in various types of criminal behaviours (property-related vs violent crime). We will do so by identifying what percentage of young offenders are victims of child abuse, and have a home address that is associated with domestic abuse incidents. Given the problem of underreporting, this will give us a lower bound for the estimate of the proportion of children with behavioural problems who live in abusive households. \textit{?}

\begin{table}[!htbp]
\caption{The long-lasting effects of domestic abuse, regressions}
  \begin{threeparttable}[t]
  \centering
       \begin{tabular}{ L{2.8cm}  C{1.8cm}  C{1.8cm}  C{3.2cm}  C{2.8cm}  C{2cm} }
    \toprule
     \textbf{Research question} & \textbf{Dataset}    & \textbf{Unit of analysis} & \textbf{Outcome variables} & \textbf{Explanatory variables} & \textbf{Model} \\
    \midrule
    How does witnessing domestic abuse as a child affect socio-economic outcomes in adulthood? & CSEW & Individual (respondent-level) & Educational attainment, employment status and health-related outcomes, including domestic abuse victimisation & Witnessing domestic abuse & Propensity score matching \\
                \midrule
                   How does experiences of past (but not recent) domestic abuse affect socio-economic outcomes in adulthood? & CSEW & Individual (respondent-level) & Current health and employment outcomes & Degree and type of domestic abuse suffered & Propensity score matching \\
                \midrule
     How does living a household with domestic abuse affect behavioural outcomes in childhood? & CSEW & Individual (respondent-level) & Bullying, involvement with gangs, school truancy, carrying knives, drugs and alcohol usage, learning difficulties, health & Degree and type of domestic abuse in household & Propensity score matching \\
                \midrule
          Are young offenders from abusive households more violent? & WMP & Young offenders (individual) & Type of offence (property-related, public order offence, violent) & Household LSOA socio-economic characteristics, abusive household or not & Multinomial logistic regression \\
     \bottomrule
  \end{tabular}
    \end{threeparttable}%
  \label{tab:addlabel}%
\end{table}%
